### 2018-03-28 analysis/data_analysis.Rmd
Added figs

---
title: "Lineage Metrics Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readr)

```

```{r, echo=FALSE}
alex_real_value_data <- read_csv("alex_real_value_data.csv")
emily_real_value_data <- read_csv("real_value_data.csv")
real_value_data <- merge(alex_real_value_data, emily_real_value_data, all = TRUE)
real_value_data$problem <- sapply(as.character(real_value_data$p),switch,"0"="Himmelblau","1"="Six-hump camel back","2"="Shubert", "6"="Composition function 2")
real_value_data$problem <- as.factor(real_value_data$problem)
real_value_data$selection <- sapply(as.character(real_value_data$selection),switch,"0"="Tournament","1"="Lexicase","2"="Eco-EA", "3"="MAP-Elites", "4"="Roulette", "5"="Drift")
real_value_data$selection <- as.factor(real_value_data$selection)
real_value_data$seed <- as.factor(real_value_data$seed)
real_value_data$ts <- as.factor(real_value_data$ts)
real_value_data$elite <- as.factor(real_value_data$elite)
real_value_data$mutation_rate <- as.factor(real_value_data$mutation_rate)
tournament_data <- subset(real_value_data, real_value_data$selection == "Tournament" & real_value_data$elite == 0)
```

```{r}
#see how many finished
summary(subset(real_value_data, real_value_data$update == 5000))

```

First, let's make sure there's nothing weird going on with fitness:

```{r, echo=FALSE}
ggplot(tournament_data) + stat_summary(aes(x=update, y=max_fitness, color=ts, fill=ts), fun.data="mean_cl_boot", geom="smooth") + theme_classic() +facet_wrap(~problem*mutation_rate, ncol=8, scales="free")
ggsave("../figs/all_tournament_fitness.png")
```

The most important thing to observe from that is probably that mutation rate seems to have a notable effect on the fitness dynamics of the Shubert function.

Onto the metrics.

## Phenotypic volatility

```{r, echo=FALSE}
ggplot(tournament_data) + stat_summary(aes(x=update, y=mean_phenotypic_volatility, color=ts, fill=ts), fun.data="mean_cl_boot", geom="smooth") + theme_classic() +facet_wrap(~problem*mutation_rate, ncol=8, scales="free")
ggsave("../figs/all_phenotypic_volatility.png")
```

Oh right, that one kind of breaks in these problems...

## Phenotype count

```{r, echo=FALSE}
ggplot(tournament_data) + stat_summary(aes(x=update, y=mean_unique_phenotypes, color=ts, fill=ts), fun.data="mean_cl_boot", geom="smooth") + theme_classic() +facet_wrap(~problem*mutation_rate, ncol=8, scales="free")
ggsave("../figs/all_phenotype_count.png")

```

These too.

## Deleterious steps

```{r, echo=FALSE}
ggplot(tournament_data) + stat_summary(aes(x=update, y=mean_deleterious_steps, color=ts, fill=ts), fun.data="mean_cl_boot", geom="smooth") + theme_classic() +facet_wrap(~problem*mutation_rate, ncol=8, scales="free")
ggsave("../figs/all_deleterious_steps.png")

```

In general, weaker selection allows more deleterious steps, as does higher mutation rate.

```{r, echo=FALSE}
ggplot(tournament_data) + stat_summary(aes(x=update, y=dominant_deleterious_steps, color=ts, fill=ts), fun.data="mean_cl_boot", geom="smooth") + theme_classic() +facet_wrap(~problem*mutation_rate, ncol=8, scales="free")
ggsave("../figs/dominant_deleterious_steps.png")

```

Looks pretty similar.

```{r, echo=FALSE}
ggplot(subset(tournament_data, tournament_data$update==5000)) + geom_point(aes(x=max_deleterious_steps, y=dominant_deleterious_steps, color=ts)) + facet_wrap(~problem*mutation_rate, ncol=8, scales="free") + theme_classic()
```

The relationship between max deleterious steps and dominant deleterious steps stops being a straight line at higher mutation rates.

# Graphs for paper

## Varying mutation rate:

### Deleterious steps:

```{r, echo=FALSE}
ggplot(subset(tournament_data, tournament_data$update==5000 & tournament_data$ts == 4)) + geom_boxplot(aes(x=mutation_rate, y=max_pairwise_distance, color=mutation_rate)) + facet_wrap(~problem, ncol=8, scales="free") + theme_classic()
```

```{r, echo=FALSE}
ggplot(subset(tournament_data, tournament_data$update==5000 & tournament_data$ts == 4)) + geom_boxplot(aes(x=mutation_rate, y=dominant_deleterious_steps, color=mutation_rate)) + facet_wrap(~problem, ncol=8, scales="free") + theme_classic()
ggsave("../figs/dom_deleterious_mutation_rate.png")
```

There's possibly an interesting story to tell about variance:

```{r, echo=FALSE}
ggplot(subset(tournament_data, tournament_data$update==5000 & tournament_data$ts == 4)) + geom_boxplot(aes(x=mutation_rate, y=variance_deleterious_steps, color=mutation_rate)) + facet_wrap(~problem, ncol=8, scales="free") + theme_classic()
ggsave("../figs/variance_deleterious_mutation_rate.png")
```

### Mutation magnitude

```{r, echo=FALSE}
ggplot(subset(tournament_data, tournament_data$update==5000 & tournament_data$ts == 4)) + geom_boxplot(aes(x=mutation_rate, y=mean_normal_mutations_on_lineage, color=mutation_rate)) + facet_wrap(~problem, ncol=8, scales="free") + theme_classic()
```

Presumably the reason that .1 is ludircously high in everything except camel-back is that on the other landscapes it's jumping between peaks. Let's look at it without .1:

```{r, echo=FALSE}
ggplot(subset(tournament_data, tournament_data$update==5000 & tournament_data$ts == 4 & as.numeric(tournament_data$mutation_rate) < 8)) + geom_boxplot(aes(x=mutation_rate, y=mean_normal_mutations_on_lineage, color=mutation_rate)) + facet_wrap(~problem, ncol=8, scales="free") + theme_classic()
```